---
title: "assignment_1"
author: "Rania Karamallah, MC Abbott, Alex Cardelle"
date: "10/20/2021"
output: 
  html_document"
    theme: readable
    toc: true
    toc_depth: 3
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question

Do communities living in neighborhoods with high walkability scores have low rates of mental health issues?

# Prior Research

[add existing research, 2 other studies is sufficient]

# Data

We will be using the following data sets:

* Walkability Score 
* Centers for Disease Control and Prevention: U.S. Chronic Disease Indicators: Mental Health 

## Load data

```{r}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
```

## total population

```{r, message=FALSE, results='hide'}
census_api_key("41cce2e06e8d30d287f5edb7186f191c45d046bc")
census <- get_decennial(geography = "state", 
                        year = 2020,
                        variables = c(pop = "P1_001N"),
                        output = "wide",
                        geometry = TRUE)
```
## Med Age

```{r, message=FALSE, results='hide'}
acs_age <- get_acs(geography = "state", 
                   variables = c(med_age_ = "B01002_001"), 
                   output = "wide")
```

## U.S. Chronic Disease Indicators: Mental Health 

```{r}
MentalHealth <- read_csv("https://raw.githubusercontent.com/mcmcmca/quant_assignment_1/main/U.S._Chronic_Disease_Indicators__Mental_Health.csv") %>%
  filter(YearStart == "2019" & YearEnd == "2019" & StratificationCategory1 == "Overall" & DataValueUnit == "Number") %>%
  select(LocationAbbr, DataValue, DataValueAlt, StratificationCategory1)
```

