---
title: "assignment_1"
author: "Rania Karamallah, MC Abbott, Alex Cardelle"
date: "11/1/2021"
output: 
  rmdformats::downcute
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question

Is there a relationship between mode of transportation to travel to work and different health indicators? How do other demographic characteristics (eg. race and age) factor in, if at all?

# Prior Research

Riggs and Sethi (2020), in examining a quantified, indexed version of "walkability", have determined that neighborhoods scoring high in this category do show increased multimodal/alternative transportation use (namely -- walking, cycling, or transit). With more specificity, Watson et. al (2020) provide a national-level overview that confirms a strong link between high walkability scores and walking itself as a mode of commuting to work throughout the urban areas of the United States. Our research question seeks to confirm these findings through selected census and walkability study data.

# Data

We will be using the following data sets:

* United States American Community Survey (2019)
* Centers for Disease Control and Prevention: PLACES U.S. Chronic Disease Indicators (2020)
* United States Environmental Protection Agency: National Walkability Index (2021)

# Variables for Analysis

Continuous
* Population Percentage with Current Mental Health Issues
* Population Percentage with Obesity
* Population Percentage with Current Asthma
* Percentage of Mode of Transportation to Work
* Walkability Index Score

Categorical
* Majority White vs. Non-White
* Majority Female vs. Not Female

## Sample Definition

Our research will focus on the scale of census tracts in Massachusetts, with a look at adult characteristics and behaviors aggregated to the level of census tracts.

## Load data

```{r}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(sf)
library(gridExtra)
```

## Mode of Transportation

Source: American Community Survey (2019)

* Mode of Transportation

```{r, message = FALSE, results = FALSE, warning = FALSE}

acs_transport <- get_acs(geography = "tract", 
                        year = 2019,
                        variables = c(transport_any = "B08301_001E",
                                      transport_walk = "B08301_019E",
                                      transport_car = "B08301_002E",
                                      transport_public = "B08301_010E",
                                      transport_bike = "B08301_018E",
                                      transport_taxi = "B08301_016E",
                                      transport_motorcycle = "B08301_017E",
                                      transport_other = "B08301_020E"),
                        state = "MA",
                        output = "wide",
                        geometry = FALSE) %>%
  mutate(pct_walk = transport_walk / transport_any) %>%
  mutate(pct_car = transport_car / transport_any) %>%
  mutate(pct_public = transport_public / transport_any) %>%
  mutate(pct_bike = transport_bike / transport_any) %>%
  mutate(pct_taxi = transport_taxi / transport_any) %>%
  mutate(pct_motorcycle = transport_motorcycle / transport_any) %>%
  mutate(pct_other = transport_other / transport_any)%>%
  select(GEOID, pct_walk, pct_car, pct_public, pct_bike, pct_taxi, pct_motorcycle, pct_other)
  
```

## Demographics

Source: American Community Survey (2019)

* Majority White vs. Non-White (categorical)
* Majority Female vs. Not Female (categorical)

```{r}

race_majority <- get_acs(geography = "tract", 
                        year = 2019,
                        variables = c( total_pop = "B02001_001",
                                       white_pop = "B02001_002"),
                        state = "MA",
                        output = "wide",
                        geometry = FALSE) %>%
  mutate(white_majority = case_when(white_popE / total_popE >= .5 ~ "Majority White",
                          white_popE / total_popE < .5 ~ "Non-Majority White",
                          TRUE ~ "unknown")) %>%
  select(GEOID,white_majority)

sex_majority <- get_acs(geography = "tract", 
                        year = 2019,
                        variables = c( total_pop = "B01001_001",
                                       female_pop = "B01001_026"),
                        state = "MA",
                        output = "wide",
                        geometry = FALSE) %>%
  mutate(female_majority = case_when(female_popE / total_popE >= .5 ~ "Female Majority",
                          female_popE / total_popE < .5 ~ "Non-Female Majority",
                          TRUE ~ "unknown")) %>%
  select(GEOID,female_majority)

```

## U.S. Chronic Disease Indicators

Source: United States Centers for Disease Control and Prevention: PLACES U.S. Chronic Disease Indicators

* Population Percentage with Current Mental Health Issues (continuous)
* Population Percentage with Obesity (continuous)
* Population Percentage with Current Asthma (continuous)

```{r}
Mental_Health <- read_csv('data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2020_release.csv') %>%
  filter(Year == 2018) %>%
  filter(Short_Question_Text == "Mental Health") %>%
  filter(StateAbbr == "MA") %>%
  rename(GEOID = LocationID) %>%
  group_by(GEOID) %>%
  rename(pct_mental = Data_Value) %>%
  select(GEOID, pct_mental, Data_Value_Unit)

Asthma <- read_csv('data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2020_release.csv') %>%
  filter(Year == 2018) %>%
  filter(Short_Question_Text == "Current Asthma") %>%
  filter(StateAbbr == "MA") %>%
  rename(GEOID = LocationID) %>%
  group_by(GEOID) %>%
  rename(pct_asthma = Data_Value) %>%
  select(GEOID, pct_asthma, Data_Value_Unit)

Obesity <- read_csv('data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2020_release.csv') %>%
  filter(Year == 2018) %>%
  filter(Short_Question_Text == "Obesity") %>%
  filter(StateAbbr == "MA") %>%
  rename(GEOID = LocationID) %>%
  group_by(GEOID) %>%
  rename(pct_obesity = Data_Value) %>%
  select(GEOID, pct_obesity, Data_Value_Unit)

```
## Walkability

Source: United States Environmental Protection Agency: National Walkability Index

* Walkability Index Score

```{r}
Walkability <- st_read(dsn = "Natl_WI.gdb")  %>%
  filter(STATEFP == "25") %>%
  select(GEOID10, NatWalkInd, TotPop) %>%
  mutate(tract = substr(GEOID10, 1, 11)) %>%
  st_set_geometry(NULL) %>%
  group_by(tract) %>%
  summarise(avg_walk_ind = weighted.mean(NatWalkInd, TotPop)) %>%
  rename(GEOID = tract)
```

# Assembling the datasets 

```{r, message = FALSE, warning = FALSE}
data <- left_join(acs_transport, Walkability) %>% 
  left_join(Asthma) %>%
  left_join(Obesity) %>%
  left_join(Mental_Health) %>%
  left_join(race_majority) %>%
  left_join(sex_majority) %>%
  select(GEOID, avg_walk_ind, pct_walk, pct_car, pct_public, pct_bike, pct_taxi, pct_motorcycle, pct_other, pct_asthma, pct_mental, pct_obesity, white_majority, female_majority)

kable(head(data))
```

# Initial Analysis of Continuous Data

```{r}
asthma_t_test <- t.test(data$pct_asthma)
obesity_t_test <- t.test(data$pct_obesity)
mental_t_test <- t.test(data$pct_mental)
walkability_t_test <- t.test(data$avg_walk_ind)
mwalk_t_test <- t.test(data$pct_walk)
mcar_t_test <- t.test(data$pct_car)
mpublic_t_test <- t.test(data$pct_public)
mbike_t_test <- t.test(data$pct_bike)
mtaxi_t_test <- t.test(data$pct_taxi)
mmotorcycle_t_test <- t.test(data$pct_motorcycle)
mother_t_test <- t.test(data$pct_other)

```

```{r}

asthma_quartiles <- quantile(data$pct_asthma, na.rm = TRUE)
obesity_quartiles <- quantile(data$pct_obesity, na.rm = TRUE)
mental_quartiles <- quantile(data$pct_mental, na.rm = TRUE)

walk_quartiles <- quantile(data$pct_walk, na.rm = TRUE)
bike_quartiles <- quantile(data$pct_bike, na.rm = TRUE)
car_quartiles <- quantile(data$pct_car, na.rm = TRUE)
taxi_quartiles <- quantile(data$pct_taxi, na.rm = TRUE)
public_quartiles <- quantile(data$pct_public, na.rm = TRUE)
motorcycle_quartiles <- quantile(data$pct_motorcycle, na.rm = TRUE)
other_quartiles <- quantile(data$pct_other, na.rm = TRUE)

```

```{r}

asthma_st_dev <- sd(data$pct_asthma, na.rm = TRUE)
obesity_st_dev <- sd(data$pct_obesity, na.rm = TRUE)
mental_st_dev <- sd(data$pct_mental, na.rm = TRUE)

walk_st_dev <- sd(data$pct_walk, na.rm = TRUE)
bike_st_dev <- sd(data$pct_bike, na.rm = TRUE)
car_st_dev <- sd(data$pct_car, na.rm = TRUE)
taxi_st_dev <- sd(data$pct_taxi, na.rm = TRUE)
public_st_dev <- sd(data$pct_public, na.rm = TRUE)
motorcycle_st_dev <- sd(data$pct_motorcycle, na.rm = TRUE)
other_st_dev <- sd(data$pct_other, na.rm = TRUE)

```


```{r}
asthma_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_asthma),
                 bins = 30)

obesity_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_obesity),
                 bins = 30) +
  scale_x_continuous(trans = "log")

mental_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_mental),
                 bins = 30)

```

```{r}
mode_summary <- tibble(
  Variable = c("Percentage of Workers Who Commute by Walking", 
               "Percentage of Workers Who Commute by Bike", 
               "Percentage of Workers Who Commute by Car",
               "Percentage of Workers Who Commute by Public Transit",
               "Percentage of Workers Who Commute by Taxi",
               "Percentage of Workers Who Commute by Motorcycle",
               "Percentage of Workers Who Commute by Other Means"),
  `Sample mean` = c((mwalk_t_test$estimate*100),
                    (mbike_t_test$estimate*100),
                    (mcar_t_test$estimate*100),
                    (mpublic_t_test$estimate*100),
                    (mtaxi_t_test$estimate*100),
                    (mmotorcycle_t_test$estimate*100),
                    (mother_t_test$estimate*100)),
  
  `Population mean (95% confidence) - low` = 
    c((mwalk_t_test$conf.int[1]*100),
      (mbike_t_test$conf.int[1]*100),
      (mcar_t_test$conf.int[1]*100),
      (mpublic_t_test$conf.int[1]*100),
      (mtaxi_t_test$conf.int[1]*100),
      (mmotorcycle_t_test$conf.int[1]*100),
      (mother_t_test$conf.int[1]*100)),
  
  `Population mean (95% confidence) - high` =
    c((mwalk_t_test$conf.int[2]*100),
      (mbike_t_test$conf.int[2]*100),
      (mcar_t_test$conf.int[2]*100),
      (mpublic_t_test$conf.int[2]*100),
      (mtaxi_t_test$conf.int[2]*100),
      (mmotorcycle_t_test$conf.int[2]*100),
      (mother_t_test$conf.int[2])*100),
  
  Median = c((mwalk_t_test$conf.int[3]*100),
      (mbike_t_test$conf.int[3]*100),
      (mcar_t_test$conf.int[3]*100),
      (mpublic_t_test$conf.int[3]*100),
      (mtaxi_t_test$conf.int[3]*100),
      (mmotorcycle_t_test$conf.int[3]*100),
      (mother_t_test$conf.int[3]*100)),
  
  `Interquartile range` = c(((walk_quartiles[4] - walk_quartiles[2])*100),
                            ((bike_quartiles[4] - bike_quartiles[2])*100),
                            ((car_quartiles[4] - car_quartiles[2])*100),
                            ((public_quartiles[4] - public_quartiles[2])*100),
                            ((taxi_quartiles[4] - taxi_quartiles[2])*100),
                            ((motorcycle_quartiles[4] - motorcycle_quartiles[2])*100),
                            ((other_quartiles[4] - other_quartiles[2])*100)),
  `Standard deviation` = c((walk_st_dev*100), (bike_st_dev*100), (car_st_dev*100), (taxi_st_dev*100), (public_st_dev*100), (motorcycle_st_dev*100), (other_st_dev*100)))

kable(mode_summary, digits = 2)

```

```{r}

pretty_asthma_hist <- asthma_hist +
  theme_bw() +
  scale_x_continuous(name = "Percentage of residents with asthma") +
  scale_y_continuous(name = "Number of tracts") +
  theme(axis.text.x = element_text(angle = 90))

pretty_obesity_hist <- obesity_hist +
  theme_bw() +
  scale_x_continuous(name = "Percentage of residents with obesity") +
  scale_y_continuous(name = "Number of tracts") +
  theme(axis.text.x = element_text(angle = 90))

pretty_mental_hist <- mental_hist +
  theme_bw() +
  scale_x_continuous(name = "Percentage of residents with poor mental health") +
  scale_y_continuous(name = "Number of tracts") +
  theme(axis.text.x = element_text(angle = 90))

grid.arrange(pretty_asthma_hist, pretty_obesity_hist, pretty_mental_hist,
             ncol = 3)

```

# Initial Analysis of Categorical Data

```{r}
pct_white <- t.test(data$white_majority == "Majority White")
pct_nonwhite <-  t.test(data$white_majority == "Non-Majority White")

pct_female <- t.test(data$female_majority == "Female Majority")
pct_notfemale <-  t.test(data$female_majority == "Non-Female Majority")

```

```{r}

cat_summary_race <- tibble(`Majority Race` = 
                              c("White",
                                "Not White"),
                            `Sample proportion` = 
                              c(pct_white$estimate * 100,
                                pct_nonwhite$estimate *100),
                            `95-percent confidence - low` = 
                              c(pct_white$conf.int[1] * 100,
                                pct_nonwhite$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_white$conf.int[2] * 100,
                                pct_nonwhite$conf.int[2] * 100))

kable(cat_summary_race, digits = 2)

```

```{r}

cat_summary_sex <- tibble(`Majority Sex` = 
                              c("Female",
                                "Not Female"),
                            `Sample proportion` = 
                              c(pct_female$estimate * 100,
                                pct_notfemale$estimate *100),
                            `95-percent confidence - low` = 
                              c(pct_female$conf.int[1] * 100,
                                pct_notfemale$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_female$conf.int[2] * 100,
                                pct_notfemale$conf.int[2] * 100))

kable(cat_summary_sex, digits = 2)

```



# Resources

Centers for Disease Control and Prevention, "PLACES: Local Data for Better Health, Census Tract Data 2020 release". 2020.
https://chronicdata.cdc.gov/500-Cities-Places/PLACES-Local-Data-for-Better-Health-Census-Tract-D/cwsq-ngmh

United States Census Bureau. American Community Survey, 5-year estimates. 2019.

Watson et al., "Associations between the National Walkability Index and walking among US Adults - National Health Interview Survey, 2015". 2020.
https://pubmed.ncbi.nlm.nih.gov/32389677/

Riggs, William Warren and Suresh Andrew Sethi, "Multimodal travel behaviour, walkability indices, and social mobility: how neighbourhood walkability, income and household characteristics guide walking, biking & transit decisions". 2020.
https://www.researchgate.net/publication/344647424_Multimodal_travel_behaviour_walkability_indices_and_social_mobility_how_neighbourhood_walkability_income_and_household_characteristics_guide_walking_biking_transit_decisions