---
title: "assignment_1"
author: "Rania Karamallah, MC Abbott, Alex Cardelle"
date: "10/20/2021"
output: 
  html_document"
    theme: readable
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question

Is there a relationship between mode of transportation and different health indicators?

# Prior Research -- UPDATE

Riggs and Sethi (2020), in examining a quantified, indexed version of "walkability", have determined that neighborhoods scoring high in this category do show increased multimodal/alternative transportation use (namely -- walking, cycling, or transit). With more specificity, Watson et. al (2020) provide a national-level overview that confirms a strong link between high walkability scores and walking itself as a mode of commuting to work throughout the urban areas of the United States. Our research question seeks to confirm these findings through selected census tracts.

# Data

We will be using the following data sets:

* United States Environmental Projection Agency: Smart Location Mapping, National Walkability Index
* United States American Community Survey: Mode of Transportation
* Centers for Disease Control and Prevention: U.S. Chronic Disease Indicators

# Variables for Analysis

* Population Percentage with Current Mental Health Issues (continuous)
* Population Percentage with Obesity (continuous)
* Population Percentage with Current Asthma (continuous)
* Mode of Transportation (continuous)
* Walkability Index Score (continuous)
* Urban vs. Rural (categorical) -- TO BE ADDED
* +1 (categorical) -- TO BE ADDED

## Load data

```{r}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
```

## Mode of Transportation

Source: American Community Survey (2019)

* Mode of Transportation (continuous)

```{r, message = FALSE, results = FALSE, warning = FALSE}

acs_transport <- get_acs(geography = "tract", 
                        year = 2019,
                        variables = c(transport_any = "B08301_001",
                                      transport_walk = "B08301_019",
                                      transport_car = "B08301_002",
                                      transport_public = "B08301_010",
                                      transport_bike = "B08301_018",
                                      transport_taxi = "B08301_016",
                                      transport_motorcycle = "B08301_018",
                                      transport_other = "B08301_020"),
                        state = "MA",
                        county = "Suffolk",
                        output = "wide",
                        geometry = FALSE) %>%
  mutate(pct_walk = transport_walk / transport_any) %>%
  mutate(pct_car = transport_car / transport_any) %>%
  mutate(pct_public = transport_public / transport_any) %>%
  mutate(pct_bike = transport_bike / transport_any) %>%
  mutate(pct_taxi = transport_taxi / transport_any) %>%
  mutate(pct_motorcycle = transport_motorcycle / transport_any) %>%
  mutate(pct_other = transport_other / transport_any)%>%
  select(GEOID, pct_walk, pct_car, pct_public, pct_bike, pct_taxi, pct_motorcycle, pct_other)
  
```

## U.S. Chronic Disease Indicators

Source: United States Center for Disease Control

* Population Percentage with Current Mental Health Issues (continuous)
* Population Percentage with Obesity (continuous)
* Population Percentage with Current Asthma (continuous)

```{r}
Mental_Health <- read_csv('data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2020_release.csv') %>%
  filter(Year == 2018) %>%
  filter(Short_Question_Text == "Mental Health") %>%
  filter(StateAbbr == "MA") %>%
  rename(GEOID = LocationID) %>%
  group_by(GEOID) %>%
  rename(pct_mental = Data_Value) %>%
  select(GEOID, pct_mental, Data_Value_Unit)

Asthma <- read_csv('data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2020_release.csv') %>%
  filter(Year == 2018) %>%
  filter(Short_Question_Text == "Current Asthma") %>%
  filter(StateAbbr == "MA") %>%
  rename(GEOID = LocationID) %>%
  group_by(GEOID) %>%
  rename(pct_asthma = Data_Value) %>%
  select(GEOID, pct_asthma, Data_Value_Unit)

Obesity <- read_csv('data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2020_release.csv') %>%
  filter(Year == 2018) %>%
  filter(Short_Question_Text == "Obesity") %>%
  filter(StateAbbr == "MA") %>%
  rename(GEOID = LocationID) %>%
  group_by(GEOID) %>%
  rename(pct_obesity = Data_Value) %>%
  select(GEOID, pct_obesity, Data_Value_Unit)

```

## Walkability

From USEPA: The National Walkability Index provides walkability scores based on a simple formula that ranks selected indicators from the Smart Location Database that have been demonstrated to affect the propensity of walk trips. Currently, the National Walkability Index dataset ranks each block group relative to all other block groups in the United States, but individuals can use downloadable data and the "how-to" document to construct an index for a smaller universe of block groups, like a state, metropolitan area, or city.

```{r}
Walkability <- read_csv("data/EPA_SmartLocationDatabase_V3_Jan_2021_Final.csv")  %>%

  filter(STATEFP == "25") %>%
  select(GEOID10, NatWalkInd)

```

## Assembling the datasets 

```{r, message = FALSE, warning = FALSE}
#data <- left_join(census, GEOID) %>% 
#  left_join(acs_transport) %>%
#  left_join(Mental_Health) %>%
#  left_join(Asthma) %>%

#kable(head(data))
```
